---
title: "Geographical tags - African Manager"
author: "Claude Grasland"
format: html
self-contained: true
---


The aim of this program is to import data from a text file and to transform them into an  annotated corpus with national and macroregional geographical units.

```{r setup1, echo = TRUE,  warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE,  warning = FALSE, message = FALSE)
library(data.table)
library(knitr)
library(quanteda)
library(ggplot2)
library(stringr)
library(dplyr)
library(tidytext)
```

# LOAD & CLEAN

## Load data

We import the data and put them in a single data.frame. Then we select the columns of interest for the corpus and rename them. 

```{r corpus_load}


df<-readRDS("../corpus_raw/fr_TUN_afrman.RDS")
# select and rename columns of interest
df$id <- 1:dim(df)[1]                   # Choose unique id
df$who <- df$media                      # Precise source
df$when <- df$date                      # Load date
df$text <- df$title                     # Load text
df$lang <- df$lang                      # Precise language
df<-df[,c("id","who","when","lang","text")]


# Order by time period
df<-df[order(df$when),]

# eliminate duplicate
df<-df[duplicated(df$text)==F,]

# eliminate non complete fields
df<-df[complete.cases(df),]

# Identify language
mylang<-"fr"

# Identify host country
mycountry<-"TUN"


```




## Clean data

The aim of this step is to clean the text an eliminate pieces of information considered as non relevant. The solution is different according to research objectives.

The aim of this step is to clean the text an eliminate pieces of information considered as non relevant. The solution is different according to research objectives.

In the case of the African Manager, it was necessary to replace "BCE" by "Beji Caïd Essebsi" because in this country BCE is always associated to the tunisian President and not to the European Central Bank as it is the case in France. 


```{r}

x<-df

# Clean encoding errors
x$text<-gsub("&apos;","'",x$text)

# Eliminate pieces of text

# Control specific ambiguities
x$text<-str_replace_all(x$text,pattern = "la BCE","la Banque Centrale Européenne")
x$text<-str_replace_all(x$text,pattern = "La BCE","La Banque Centrale Européenne")
x$text<-str_replace_all(x$text,pattern = "(BCE)","")
x$text<-str_replace_all(x$text,pattern = "BCE","Beji Caïd Essebsi")

df<-x
rm(x)
```


## Transform in quanteda corpus

We transform the datafram into quanteda corpus object in order to obtain informations on the number of tokens and eventually eliminate short texts. 

### Transform in quanteda corpus

```{r}
qd<-corpus(df,docid_field = "id",text_field = "text")
```


### Control the number of tokens

```{r tokens}

# Compute number of tokens
qd$nbt<-ntoken(qd)

# Choose news with minimum / maximum number of tokens
qd<-corpus_subset(qd, nbt<100)
qd<-corpus_subset(qd, nbt>4)


```




# TAG

## Preparation of data


### Load dictonary

We loading the last version of the Imageun dictionary and we extract our target language 

```{r load_dict}
# Load general dictionary
dict<-read.csv2("../dict/dict_final.csv")


# select language
dict <- dict[dict$lang==mylang,-1]


# Visualize
head(dict)
```


### Load tagging function

```{r func_annotate}
extract_tags <- function(qd = qd,                       # the corpus of interest
                         lang = "en",                   # the language to be used
                         dict = dict,                   # the dictionary of target 
                         code = "id" ,                  # variable used for coding
                         tagsname = "tags",             # name of the tags column
                         split  = c("'","’","-",":"),   # split list (option)
                         tolow = FALSE                  # transform lower case (option)
                         )
{ 


  
# Tokenize  
x<-as.character(qd)

### Split (optional)
if(length(split) > 0) { reg<-paste(split, collapse = '|')
                       x <- gsub(reg," ",x)} 
### To lower (optional)
if(tolow) { x <- tolower(x)} 

### Tokenize
toks<-tokens(x)



  
# Load dictionaries and create compounds

### Prepare dictionary
dict<-dict[dict$lang==lang & is.na(dict$alias)==F,]
labels <-dict$alias
if(length(split) > 0) { reg<-paste(split, collapse = '|')
                       labels<- gsub(reg," ",labels)}  
if(tolow)       {labels <- tolower(labels)}  
toks<-tokens_compound(toks,pattern=phrase(labels))
  
 # create quanteda dictionary
keys <-gsub(" ","_",labels)
qd_dict<-as.list(keys)
names(qd_dict)<-dict[[code]]
qd_dict<-dictionary(qd_dict,tolower = FALSE)

# Identify geo tags (states or reg or org ...)
toks_tags <- tokens_lookup(toks, qd_dict, case_insensitive = F)
toks_tags <- lapply(toks_tags, unique)
toks_tags<-as.tokens(toks_tags)
list_tags<-function(x){res<-paste(x, collapse=' ')}
docvars(qd)[[tagsname]]<-as.character(lapply(toks_tags,FUN=list_tags))
docvars(qd)[[paste("nb",tagsname,sep="")]]<-ntoken(toks_tags)



# Export results
return(qd)
 }
```



## Geographical annotation

### Annotate all entities

In a first step, we annotate all geographic entities together in order to benefit from the cross-definition of their respective compounds. We will separate them by subcategories in a second step. Each language is done separately.


```{r annotate}



qd <- extract_tags (qd = qd,
                     lang=mylang,
                     dict = dict,
                     code = "code",
                     tagsname = "geo",
                     split = c("'","’","-",":"),
                     tolow = FALSE)


```



### Extract states codes

```{r extract_states}
# List of entities
ent <-dict %>% filter(duplicated(code)==F) %>% select(code,code_name)

# selection of states
state<-ent$code[substr(ent$code,1,3)=="STA"]

# Extract states codes
test <- paste(state, collapse="|")
x<-as.character(lapply(str_extract_all(qd$geo,paste(test, collapse = '|')), paste,collapse=" "))

# Keep only ISO3 code
x<-gsub("STA_NAM_","",x)
x<-gsub("STA_CAP_","",x)
x<-gsub("STA_HEA_","",x)
x<-gsub("_.+","",x)

# Eliminate host country (optional)
#x<-gsub(mycountry,"",x)

# Eliminate duplicatd states
y<-tokens(x)
y<-lapply(y, unique)



list_tags<-function(x){res<-paste(x, collapse=' ')}
qd$states<-as.character(lapply(y,FUN=list_tags))
qd$nbstates<-ntoken(qd$states)

summary(qd,3)
```

### check news with maximum state number

```{r check_states_news}
table(qd$nbstates)
check<-corpus_subset(qd,nbstates>3)
x<-data.frame(who=check$who,when = check$when,text=as.character(check),states=check$states,nbstates=check$nbstates)
x<-x[order(x$nbstates,decreasing = T),]
kable(head(x))
```




### Extract world region codes

We do not distinguish so-called "geographical" regions (like "Europe") and "political" regions (like "European Union") and put them in the same catagory of world regions i.e. first level of organization under the world level and/or first level of agregation above state level.


```{r extract_regions}
# Extract codes of regions
region<-ent$code[substr(ent$code,1,3)=="REG"]

# Reorder list of regions with same beginning
region<-region[order(-nchar(region))]
test <- paste(region, collapse="|")
x<-as.character(lapply(str_extract_all(qd$geo,paste(test, collapse = '|')), paste,collapse=" "))
x<-gsub("REG_","",x)

# collapse EU variant (optional)
x<-gsub("ORG_EUR_EU.+","ORG_EUR_EU",x)


y<-tokens(x)
y<-lapply(y, unique)
list_tags<-function(x){res<-paste(x, collapse=' ')}
qd$regions<-as.character(lapply(y,FUN=list_tags))
qd$nbregions<-ntoken(qd$regions)
table(qd$nbregions)

```

### Check news with maximum number of world regions

```{r check_regions_news}
table(qd$nbregions)
check<-corpus_subset(qd,nbregions>1)
x<-data.frame(who=check$who,when = check$when,text=as.character(check),regions=check$regions,nbregions=check$nbregions)
x<-x[order(x$nbregions,decreasing = T),]
kable(head(x))
```

### Check news with mixtures of states and regions

```{r check_states_regions_news}

check<-corpus_subset(qd,nbregions>0 & nbstates >1)
x<-data.frame(who=check$who,when = check$when,text=as.character(check),geo=check$geo,nbstates=check$nbstates, nbregions = check$nbregions)
x<-x[order(x$nbstates*x$nbregions,decreasing = T),]
kable(head(x))
```


### Save geographically anotated corpus

```{r}
td<-tidy(qd)
saveRDS(td,"../corpus_geo/fr_TUN_afrman_geo.RDS")
```


### Extract sample for test (optional)

We extract 200 news :

- 50 news with one foreign state
- 50 news with one macroregion
- 50 news with one foreign state and one macroregion
- 50 news with no foreign states and no macroregions

```{r, eval=FALSE}

set.seed(42)
# (E) Prepare sample test
k<-50
##  news with one single foreign country
test1<-corpus_subset(qd, nbstates==1 & substr(states,1,3) !=mycountry) %>% 
        corpus_sample(k) %>% tidy()

## news with one single macro-region
test2<-corpus_subset(qd, nbregions==1) %>% 
  corpus_sample(k) %>% tidy()

## news with one foreign country and one macro region
test3<-corpus_subset(qd, nbstates==1 & substr(states,1,3) !=mycountry & nbregions==1 ) %>% 
  corpus_sample(k) %>% tidy()

## news with no countries and no macroregions
test4<-corpus_subset(qd, nbgeo==0) %>% 
  corpus_sample(k) %>% tidy()

test<-rbind(test1,test2,test3,test4)

res<-data.frame(text=test$text, 
                tags=test$geo, 
                S=as.numeric(test$nbstates>0),
                Spos=0,
                Sneg=0,
                R=as.numeric(test$nbregions>0),
                Rpos=0,
                Rneg=0,
                Comment="ok")


write.table(res,"../test/fr_TUN_afrman_geo_test.csv", row.names = F, sep=";")
```





